Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-03-11 paper/paper.tex
update polynomial

\subsection{$k$-Coloring}
Let us use the 3-coloring problem defined on vertices as an example.
For a vertex $v$, we define the degrees of freedom $c_v\in\{0,1,2\}$ and a vertex tensor labelled by it as
\begin{equation}
    W(v) = \left(\begin{matrix}
        r_v\\
        g_v\\
        b_v
    \end{matrix}\right),
\end{equation}
where $r_v$, $g_v$ and $b_v$ are colors used to label the configurations.
For an edge $(u, v)$, we define an edge tensor as a matrix labelled by $(c_u, c_v)$ to specify the constraint
\begin{equation}
    B(x) = \left(\begin{matrix}
        0 & x & x\\
        x & 0 & x\\
        x & x & 0
    \end{matrix}\right).
\end{equation}
The number of possible colorings can be obtained by contracting this tensor network by setting $r_v, g_v$ and $b_v$ to $1$.
By designing generic types as tensor elements, one can get other properties.
Similarly, one can also define the $k$-coloring problem on edges by switching the roles of edges and vertices.

\subsection{Max-Cut problem}
The Max-Cut problem is also known as the boolean spin glass problem.
For a vertex $v\in V$, we define a boolean degree of freedom $s_v\in\{0, 1\}$.
Then the Max-Cut problem can be encoded to tensor networks by mapping an edge $(i,j)\in E$ to an edge matrix labelled by $s_is_j$
\begin{equation}
    B(x_{\langle i, j\rangle}) = \left(\begin{matrix}
        1 & x_{\langle i, j\rangle}\\
        x_{\langle i, j\rangle} & 1
    \end{matrix}\right),
\end{equation}
where variable $x_{\langle i, j\rangle}$ represents a cut on edge $(i, j)$ or a domain wall of an Ising spin glass.
Similar to other problems, we can define a polynomial with edge variables by setting $x_{\langle i, j\rangle} = x$,
where its $k$th coefficient is two times the number of configurations of cut size $k$.

\subsection{Dominating Set}
Let $G=(V,E)$ be the target graph that we want to solve.
The tensor network representation map a vertex $v\in V$ to a boolean degree of freedom $s_v\in\{0, 1\}$.
We defined the restriction on a vertex and its neighbouring vertices $N(v)$:
\begin{equation}
T(x_v)_{s_1,s_2,\ldots,s_{|N(v)|},s_v} = \begin{cases}
    0 & s_1=s_2=\ldots=s_{|N(v)|}=s_v=0,\\
    1 & s_v=0,\\
    x_v^{w_v} & \text{otherwise},
\end{cases}
\end{equation}
where $w_v$ is the weight of vertex $v$.
This tensor means if both $v$ and its neighbouring vertices are not in $D$, i.e., $s_1=s_2=\ldots=s_{|N(v)|}=s_v=0$,
this configuration is forbidden because $v$ is not adjacent to any member in the set.
otherwise, if $v$ is in $D$, it has a contribution $x_v^{w_v}$ to the final result.

\subsection{Satisfiability Problem}
We can contruct a satisfiability problem to solve the above problem.
To generate a tensor network, we map a boolean variable $x$ and its negation $\neg x$ to a degree of freedom (label) $s_x \in \{0, 1\}$,
where 0 stands for variable $x$ having value \texttt{false} while 1 stands for having value \texttt{true}.
Then we map a clause to a tensor. For example, a clause $\neg x \vee y \vee \neg z$ can be mapped to a tensor labeled by $(s_x, s_y, s_z)$.
\begin{equation}
C = \left(\begin{matrix}
\left(\begin{matrix}
x & x \\
x & x
\end{matrix}\right) \\
\left(\begin{matrix}
x & x \\
1 & x
\end{matrix}\right)
\end{matrix}\right).
\end{equation}
There is only one entry $(s_x, s_y, s_z) = (1, 0, 1)$ that makes this clause unsatisfied.
If we contract this tensor network, we will get a multiplicative factor $x$ whenever there is a clause satisfied.

>>>>

\subsection{$k$-Coloring}
Let us use the 3-coloring problem defined on vertices as an example.
For a vertex $v$, we define the degrees of freedom $c_v\in\{0,1,2\}$ and a vertex tensor labelled by it as
\begin{equation}
    W(v) = \left(\begin{matrix}
        r_v\\
        g_v\\
        b_v
    \end{matrix}\right),
\end{equation}
where $r_v$, $g_v$ and $b_v$ are colors used to label the configurations.
For an edge $(u, v)$, we define an edge tensor as a matrix labelled by $(c_u, c_v)$ to specify the constraint
\begin{equation}
    B(x) = \left(\begin{matrix}
        0 & x & x\\
        x & 0 & x\\
        x & x & 0
    \end{matrix}\right).
\end{equation}
The number of possible colorings can be obtained by contracting this tensor network by setting $r_v, g_v$ and $b_v$ to $1$.
By designing generic types as tensor elements, one can get other properties.
Similarly, one can also define the $k$-coloring problem on edges by switching the roles of edges and vertices.
Let $x_{\langle i,j\rangle} = x$, we have a graph polynomial,
where its $k$th coefficient is the number of coloring with $k$ bonds are satisfied.
If a graph is colorable, the maximum order of this polynomial should be equal to the number of edges in this graph.

\subsection{Max-Cut problem}
The Max-Cut problem is also known as the boolean spin glass problem.
For a vertex $v\in V$, we define a boolean degree of freedom $s_v\in\{0, 1\}$.
Then the Max-Cut problem can be encoded to tensor networks by mapping an edge $(i,j)\in E$ to an edge matrix labelled by $s_is_j$
\begin{equation}
    B(x_{\langle i, j\rangle}) = \left(\begin{matrix}
        1 & x_{\langle i, j\rangle}\\
        x_{\langle i, j\rangle} & 1
    \end{matrix}\right),
\end{equation}
where variable $x_{\langle i, j\rangle}$ represents a cut on edge $(i, j)$ or a domain wall of an Ising spin glass.
Let $x_{\langle i,j\rangle} = x$, we have a graph polynomial similar to previous ones,
where its $k$th coefficient is two times the number of configurations of cut size $k$.

\subsection{Dominating Set}
Let $G=(V,E)$ be the target graph that we want to solve.
The tensor network representation map a vertex $v\in V$ to a boolean degree of freedom $s_v\in\{0, 1\}$.
We defined the restriction on a vertex and its neighbouring vertices $N(v)$:
\begin{equation}
T(x_v)_{s_1,s_2,\ldots,s_{|N(v)|},s_v} = \begin{cases}
    0 & s_1=s_2=\ldots=s_{|N(v)|}=s_v=0,\\
    1 & s_v=0,\\
    x_v^{w_v} & \text{otherwise},
\end{cases}
\end{equation}
where $w_v$ is the weight of vertex $v$.
This tensor means if both $v$ and its neighbouring vertices are not in $D$, i.e., $s_1=s_2=\ldots=s_{|N(v)|}=s_v=0$,
this configuration is forbidden because $v$ is not adjacent to any member in the set.
otherwise, if $v$ is in $D$, it has a contribution $x_v^{w_v}$ to the final result.
The graph polynomial for the dominating set problem is known as the domination polynomial~\cite{Alikhani2009},
which is defined as
\begin{equation}
D(G, x) = \sum_{k=0}^{\gamma(G)} d_k x^k,
\end{equation}
where $d_k$ is the number of dominating sets of size $k$ in graph $G=(V, E)$.

\subsection{Satisfiability Problem}
We can contruct a satisfiability problem to solve the above problem.
To generate a tensor network, we map a boolean variable $a$ and its negation $\neg a$ to a degree of freedom (label) $s_a \in \{0, 1\}$,
where 0 stands for variable $a$ having value \texttt{false} while 1 stands for having value \texttt{true}.
Then we map a clause to a tensor. For example, a clause $\neg a \vee b \vee \neg c$ can be mapped to a tensor labeled by $(s_a, s_b, s_c)$.
\begin{equation}
C = \left(\begin{matrix}
\left(\begin{matrix}
x & x_{b} \\
x_a & x_{ab}
\end{matrix}\right) \\
\left(\begin{matrix}
x_{c} & x_{bc} \\
1 & x_{abc}
\end{matrix}\right)
\end{matrix}\right).
\end{equation}
There is only one entry $(s_x, s_y, s_z) = (1, 0, 1)$ that makes this clause unsatisfied.
If we contract this tensor network, we will get a multiplicative factor $x$ whenever there is a clause satisfied.
Let $x_{ab\ldots} = x$, then we can get a polynomial that its $k$-th coefficients gives us the number of assignments with $k$ clauses satisfied.
